Cluster Difference Imaging Photometric Survey (CDIPS) II. TOI$\,$837: A Validated Saturn-Sized Planet in IC~2602

### 2020-07-13 paper/authors.tex
update manuscript

%PENDING (MPIA)
\author{P. Sarkis} %Paula sarkis@mpia.de PI of FEROS time
\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, Heidelberg 69117, Germany }

@@

\author{W. Marie-Sainte} % Wenceslas
\affiliation{Institut Paul \’Emile Victor, Concordia Station, Antarctica}

@@

\author{T. Henning} %Thomas henning@mpia.de FEROS, WINE board
\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, Heidelberg 69117, Germany }

\author[0000-0002-9158-7315]{R. Brahm} %rafael.brahm@uai.cl coordination, WINE board
\affiliation{Facultad de Ingeniería y Ciencias, Universidad Adolfo Ib\'a\~nez, Av.\ Diagonal las Torres 2640, Pe\~nalol\'en, Santiago, Chile}
\affiliation{Millennium Institute for Astrophysics, Chile}

@@

\author[0000-0002-5389-3944]{A. Jord\'an} %andres.jordan@gmail.com WINE board
\affiliation{Facultad de Ingeniería y Ciencias, Universidad Adolfo Ib\'a\~nez, Av.\ Diagonal las Torres 2640, Pe\~nalol\'en, Santiago, Chile}
\affiliation{Millennium Institute for Astrophysics, Chile}

@@

\author{T. Trifonov} %Trifon trifonov@mpia.de FEROS observations
\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, Heidelberg 69117, Germany }

>>>>

%PENDING (MPIA)
\author{P. Sarkis} %Paula sarkis@mpia.de PI of FEROS time
\affiliation{Max-Planck-Institut f\"{u}r Astronomie, K\"onigstuhl 17, Heidelberg 69117, Germany }

@@

\author{W. Marie-Sainte} % Wenceslas
\affiliation{Institut Paul \'{E}mile Victor, Concordia Station, Antarctica}

@@

\author{T. Henning} %Thomas henning@mpia.de FEROS, WINE board
\affiliation{Max-Planck-Institut f\"{u}r Astronomie, K\"onigstuhl 17, Heidelberg 69117, Germany }

\author[0000-0002-9158-7315]{R. Brahm} %rafael.brahm@uai.cl coordination, WINE board
\affiliation{Facultad de Ingenier\'{i}a y Ciencias, Universidad Adolfo Ib\'a\~nez, Av.\ Diagonal las Torres 2640, Pe\~nalol\'en, Santiago, Chile}
\affiliation{Millennium Institute for Astrophysics, Chile}

@@

\author[0000-0002-5389-3944]{A. Jord\'an} %andres.jordan@gmail.com WINE board
\affiliation{Facultad de Ingenier\'{i}a y Ciencias, Universidad Adolfo Ib\'a\~nez, Av.\ Diagonal las Torres 2640, Pe\~nalol\'en, Santiago, Chile}
\affiliation{Millennium Institute for Astrophysics, Chile}

@@

\author{T. Trifonov} %Trifon trifonov@mpia.de FEROS observations
\affiliation{Max-Planck-Institut f\"{u}r Astronomie, K\"onigstuhl 17, Heidelberg 69117, Germany }

### 2020-07-13 paper/ms.tex
update manuscript

\suppressAffiliations
%\NewPageAfterKeywords
\input{authors.tex}

\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA TESS mission, the ESA Gaia mission, photometry from El Sauce
  and ASTEP400, and spectroscopy from CHIRON, FEROS, and Veloce.  We
  find that TOI$\,$837 is a $T=9.9$ G0/F9 dwarf in the southern open
  cluster IC 2602.  The star and planet are therefore $35^{+11}_{-5}$
  million years old.  The planet is warm ($P = 8.3\,{\rm d}$) and
  slightly smaller than Jupiter.  Combining the transit photometry
  with a prior on the stellar parameters derived from the cluster
  color-magnitude diagram, we find that the transits are grazing with
  $b=0.95\pm0.01$ and $R_{\rm p} = 0.80^{+0.04}_{-0.03} \,R_{\rm
  Jup}$.  From radial velocity monitoring, we limit the minimum mass
  to less than 1.20 $M_{\rm Jup}$ (3$\sigma$).  Grazing transits are a
  cause for concern, as they are usually indicative of astrophysical
  false positive scenarios.  Our follow-up data show that such
  scenarios are highly unlikely.  Our combined multi-color photometry,
  high-resolution imaging, and radial velocities rule out hierarchical
  eclipsing binary scenarios.  Background eclipsing binary scenarios,
  though limited by speckle imaging, formally remain a 0.2\%
  possibility.  TOI 837.01 is therefore a validated adolescent
  exoplanet.  Further observations of its stellar obliquity,
  atmosphere, and mass should confirm its planetary nature beyond
  doubt, and could improve our understanding of how the physical and
  orbital properties of exoplanets change in time.
\end{abstract}

@@

A number of other processes are expected to shape young exoplanets.
After accreting, planets with gaseous envelopes are thought to cool
and contract, and their atmospheres are expected to undergo
photoevaporation \citep[{\it
e.g.},][]{Fortney_et_al_2007,Owen_Wu_2013,Fulton_et_al_2017}.  The
relative importance of contraction and photoevaporation is set by the
planetary mass, as well as the radiation environment.  The
effectiveness of photoevaporation can be inferred from observations of
planetary winds in the metastable 1083$\,$nm He line
\citep{spake_helium_2018,oklopcic_new_2018,mansfield_detection_2018}.

@@

Finding and studying systems undergoing these evolutionary changes is
a major goal in contemporary exoplanet research.  To identify stars
younger than say 1$\,$Gyr, a number of direct and indirect methods are
available \citep{soderblom_ages_2010}.  The traditional approach is to
isochronally age-date coeval groups of stars, hereafter ``clusters''
\citep[{\it
e.g.},][]{lada_embedded_2003,zuckerman_young_2004,krumholz_star_2019}.
Young field stars can also be identified using age indicators such as
stellar rotation periods, the abundance of photospheric lithium, and
chromospheric diagnostics such as calcium emission and broadband UV
emission.  Studies by for instance \citet{david_discovery_2018} and
G{.}~Zhou et al. (2020, submitted) have combined these methods for
individual field stars.  Many of the latter methods were summarized by
\citet{mamajek_improved_2008}, and have since been calibrated by {\it
e.g.},
\citet{irwin_rotational_2009,barnes_color-period_2015,meibom_spin-down_2015,angus_calibrating_2015}
and \citet{curtis_tess_2019} for stellar rotation,
\citet{zerjal_chromospherically_2017} for chromospheric activity, and
{\it e.g.}, \citet{elliott_search_2016,berger_identifying_2018} and
\citet{zerjal_galah_2019} for lithium abundances.

@@

@@

@@

@@

@@

As we will discuss in Section~\ref{subsec:starparams}, we measured the
star's radius by combining the spectroscopic effective temperature
with a broadband photometry SED fit.  We would expect, combining our
$R_\star$ and $P_{\rm rot}$ measurements, that the equatorial velocity
$v$ of the star would be $17.67 \pm 0.32 \,{\rm km\,s}^{-1}$.  Our
spectroscopically measured $v\sin i$ from CHIRON, $17.48 \pm 0.15
\,{\rm km\,s}^{-1}$ agrees reasonably well with this expectation.

@@

The maximum amplitude of the Rossiter-McLaughlin anomaly is
\citep{gaudi_prospects_2007}
\begin{align}
  \Delta V_{\rm RM} &\approx f_{\rm LD} \cdot \delta \cdot v\sin i \cdot \sqrt{1-b^2}
  \approx 14\,{\rm m\,s}^{-1},
\end{align}
for
\begin{align}
  f_{\rm LD} &= 1 - c_1 ( 1 - \mu ) - c_2 (1 - \mu )^2,
\end{align}
where $\mu \approx (1 - b^2)^{1/2}$, and for calculation purposes we
assumed $b=0.95$ and used stellar and transit parameters from
Tables~\ref{tab:starparams} and~\ref{tab:posterior}.  Although
challenging, for a 1.9$\,$hr transit of a $V=10.6$ star, a detection
could be achieved with modern spectrographs.  The next viable total
transit windows from Chile occur in January and February of 2021;
there are also a few visible per season from other southern locations.
The most precise available ephemeris, found from our joint fit of the
TESS and ground-based photometry, is as follows.
\begin{align}
  t_0\ [{\rm BJD_{TDB}}] &= 2458574.2725293 \pm 0.0005870 \\
  P\ [{\rm d}] &= 8.3248764 \pm 0.0000155 \\
  T_{14}\ [{\rm hr}] &= 1.86 \pm 0.04.
\end{align}

@@

\clearpage

@@

% \clearpage
%\clearpage

\input{ic2602_ages.tex}
\input{starparams.tex}
%\clearpage
\input{model_posterior_table.tex}
\clearpage
\input{TOI837_phot_table.tex}
\input{TOI837_rv_table.tex}
%\clearpage

%\clearpage
\bibliographystyle{yahapj}                            
\bibliography{bibliography}

>>>>

%\suppressAffiliations
%\NewPageAfterKeywords
\input{authors.tex}

\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA TESS mission, the ESA Gaia mission, ground-based photometry from El Sauce
  and ASTEP400, and spectroscopy from CHIRON, FEROS, and Veloce.  We
  find that TOI$\,$837 is a $T=9.9$ G0/F9 dwarf in the southern open
  cluster IC 2602.  The star and planet are therefore $35^{+11}_{-5}$
  million years old. Combining the transit photometry
  with a prior on the stellar parameters derived from the cluster
  color-magnitude diagram, we find that the 
  planet is warm ($P = 8.3\,{\rm d}$) and
  slightly smaller than Jupiter ($R_{\rm p} = 0.80^{+0.09}_{-0.03} \,R_{\rm Jup}$). 
  The transits are either grazing or nearly grazing.  From radial velocity
  monitoring, we limit the minimum mass to less than 1.20 $M_{\rm
  Jup}$ (3$\sigma$).  Grazing transits are a cause for concern, as
  they are usually indicative of astrophysical false positive
  scenarios.  Our follow-up data show that such scenarios are highly
  unlikely.  Our combined multi-color photometry, high-resolution
  imaging, and radial velocities rule out hierarchical eclipsing
  binary scenarios.  Background eclipsing binary scenarios, though
  limited by speckle imaging, formally remain a 0.2\% possibility.
  TOI 837.01 is therefore a validated adolescent exoplanet.  Further
  observations of its stellar obliquity, atmosphere, and mass should
  confirm its planetary nature beyond doubt, and could improve our
  understanding of how the physical and orbital properties of
  exoplanets change in time.
\end{abstract}

@@

A number of other processes are expected to shape young exoplanets.
After accreting, planets with gaseous envelopes are thought to cool
and contract, and their atmospheres are expected to undergo
photoevaporation \citep[{\it
e.g.},][]{Fortney_et_al_2007,Owen_Wu_2013,Fulton_et_al_2017,gupta_sculpting_2019,gupta_signatures_2020}.
The relative importance of contraction and photoevaporation is set by
the planetary mass, as well as the radiation environment.  The
effectiveness of photoevaporation can be inferred from observations of
planetary winds in the metastable 1083$\,$nm He line
\citep{spake_helium_2018,oklopcic_new_2018,mansfield_detection_2018}.

@@

Finding and studying systems undergoing these evolutionary changes is
a major goal in contemporary exoplanet research.  To identify stars
younger than say 1$\,$Gyr, a number of direct and indirect methods are
available \citep{soderblom_ages_2010}.  The traditional approach is to
isochronally age-date coeval groups of stars, hereafter ``clusters''
\citep[{\it
e.g.},][]{lada_embedded_2003,zuckerman_young_2004,krumholz_star_2019}.
Young field stars can also be identified
isochronally, provided that they are sufficiently massive \citep{berger_gaia-kepler_2020}. 
Other age indicators include
stellar rotation periods, the abundance of photospheric lithium, and
chromospheric diagnostics such as calcium emission and broadband UV
emission.  Studies by for instance \citet{david_discovery_2018} and
G{.}~Zhou et al. (2020, submitted) have combined these methods to age-date
individual field stars hosting transiting planets.  Many of these latter methods were summarized by
\citet{mamajek_improved_2008}, and have since been calibrated by {\it
e.g.},
\citet{irwin_rotational_2009,barnes_color-period_2015,meibom_spin-down_2015,angus_calibrating_2015}
and \citet{curtis_tess_2019} for stellar rotation,
\citet{zerjal_chromospherically_2017} for chromospheric activity, and
{\it e.g.}, \citet{elliott_search_2016,berger_identifying_2018} and
\citet{zerjal_galah_2019} for lithium abundances.

@@

\input{TOI837_phot_table.tex}

@@

\input{TOI837_rv_table.tex}

@@

\input{ic2602_ages.tex}

@@

\input{starparams.tex}

@@

As we will discuss in Section~\ref{subsec:starparams}, we measured the
star's radius by combining the spectroscopic effective temperature
with a broadband photometry SED fit.  We would expect, combining our
$R_\star$ and $P_{\rm rot}$ measurements, that the equatorial velocity
$v$ of the star would be $17.67 \pm 0.32 \,{\rm km\,s}^{-1}$.  Our
spectroscopically measured $v\sin i$ from CHIRON, $16.2 \pm 1.1
\,{\rm km\,s}^{-1}$ agrees reasonably well with this expectation.

@@

The maximum amplitude of the Rossiter-McLaughlin anomaly is
\citep{gaudi_prospects_2007}
\begin{align}
  \Delta V_{\rm RM} &\approx f_{\rm LD} \cdot \delta \cdot v\sin i \cdot \sqrt{1-b^2}
  \approx 14\,{\rm m\,s}^{-1},
\end{align}
for
\begin{align}
  f_{\rm LD} &= 1 - c_1 ( 1 - \mu ) - c_2 (1 - \mu )^2,
\end{align}
where $\mu \approx (1 - b^2)^{1/2}$, and for calculation purposes we
assumed $b=0.95$ and used stellar and transit parameters from
Tables~\ref{tab:starparams} and~\ref{tab:posterior}.  Although
challenging, for a 1.9$\,$hr transit of a $V=10.6$ star, a detection
could be achieved with modern spectrographs.  The next viable total
transit windows from Chile occur in January and February of 2021;
there are also a few visible per season from other southern locations.
The most precise available ephemeris, found from our joint fit of the
TESS and ground-based photometry, is as follows.
\begin{align}
  t_0\ [{\rm BJD_{TDB}}] &= 2458574.272529 \pm 0.000587 \nonumber \\
  P\ [{\rm d}] &= 8.3248764 \pm 0.0000155 \nonumber \\
  T_{14}\ [{\rm hr}] &= 1.86 \pm 0.04.
\end{align}

@@

% \clearpage

@@

% \input{ic2602_ages.tex}
% \input{starparams.tex}
% \input{model_posterior_table.tex}
% \input{TOI837_phot_table.tex}
% \input{TOI837_rv_table.tex}

\clearpage
\bibliographystyle{yahapj}                            
\bibliography{bibliography}
